Automating Verification of MPC Security

### 2023-06-28 popl24/composition.tex
checkpoint

\section{Compositional Properties: Garbled $\NIMO$}
\label{section-composition}

Using YGC as a case study, in this Section we will discuss
compositional properties that can be automatically verified on small
program components using techniques developed in Section
\ref{section-bruteforce}. These properties compose to yield $\NIMO$
and scale to arbitrary programs. We will express these properties
manually as pre-~and post-conditions on components, so our
general approach is in a proof assistance style. The properties
we use are refinements of $\NIMO$, and the overall technique is
enabled by the intensional character of our program distribution
formulation.

@@

\subsection{Verification of Pre-~and Post-Conditions}

A basic intuition about YGC is that garbled gates considered
individually are ``nearly'' passive secure, modulo the feature that
outputs of individual logic gates in circuits are encrypted. That is,
each gate enjoys an encrypted variant of $\NIMO$ such that the relevant
output is not in the clear but correlated flips sampled for the gate's
output wire label by the garbler, both key and pointer bits. We call
this \emph{Garbled $\NIMO$}, or $\gNIMO$. In the definition of
$\gNIMO$ we consider whether a given value $\itv$ correlates
positively or negatively to the flip chosen for a wire label, which
can be either $1$ or $0$ uniformly- recall that
Definition \ref{def-progd} posits a variable $\itv$ to refer to the
expression $\be$ when considering the distribution
$\progd(\prog,\be)$.
\begin{definition}[Garbled $\NIMO$]
  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} =
  \be_{\ttt{p}}\}, g)$ with $\iov(\prog) = S \cup V$ iff for all $H$,
  $C$ with $|C| \le |H|$, for all $\store \in \mems(S_C \cup V_C)$,
  for both $w \in \{\ttt{p},\ttt{k}\}$, and for all $\stores$ with
  domain $X = S_C \cup V_C \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$
  such that:
  $$
  \stores = \{ \store' \mid \store \subset \store' \wedge 
  \store'(\itv) = \store'(\flip{2}{\ttt{gate:}g.w}) \} 
  $$
  or
  $$
  \stores = \{ \store' \mid \store \subset \store' \wedge 
  \store'(\itv) = \neg\store'(\flip{2}{\ttt{gate:}g.w}) \} 
  $$
  %where for any $\beta$:
  %$$\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \beta \}$$
  %or
  %$$
  %\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \neg\beta \}
  %$$
  we have:
  $$
  \margd{(\condd{\progd(\prog,\be_w)}{\stores_{X-{V_C}}})}{S_H} =
  \margd{(\condd{\progd(\prog,\be_w)}{\stores})}{S_H}
  $$
\end{definition}

\subsubsection{Verification Witnesses}

\cnote{Lemmas \ref{lemma-scope} and \ref{lemma-gnimo} in this Section may need some
refinement but I think the general argument is reasonably developed.}

Our aim is to demonstrated that $\gNIMO$ is preserved by gates within circuits,
where preservation is expressed via pre-~and post-conditions. We can use
techniques in Section \ref{section-bruteforce} to automatically verify postconditions
for components, using witness programs that satisfy preconditions prior to
component execution. For example, we can automatically verify that $\ttt{encode}$
generates a program and wire values the enjoy $\gNIMO$. Then, to verify
correctness of $\ttt{and}$ gate construction and evaluation- i.e., that it
preserves $\gNIMO$- we can verify
that the $\minifed$ program and wire value generated by the following program
$\metaprot$ program satisfies $\gNIMO$:
\begin{verbatimtab}
  let secrets = encode(s1,s2) in
  andgate(0,s1,s2); evalgate(secrets.wv1,secrets.wv2)
\end{verbatimtab}
However, we need to establish that $\ttt{encode(s1,s2)}$ is an
adequate witness, in that it ensures that automated verification generalizes to
arbitrary circuit contexts.

An important part of this argument is probabilistic independence,
and in related reasoning we can leverage prior work on probabilistic
separation logic \cite{barthe2019probabilistic}. We will use an
analog of their separation formula, which we define in our setting as follows:
\begin{definition}
  We write $(\margd{\pdf{X}}{Y}) * (\margd{\pdf{X}}{X-Y})$ iff
  $\pdf{X}(\store) = \pdf{X}(\store_{Y}) * \pdf{X}(\store_{X-Y})$
  for all $\store \in \mems(X)$.
\end{definition}

\begin{definition}
  We say that $\prog$ \emph{covers} $X$ iff
  $\progd(\prog)(\store) > 0$ for all $\store\in\mems(X)$.
\end{definition}

We aim to show that gates are probabilistically independent of each
other (i.e., are separated), and this independence supports
correctness-preserving composition. A critical property is that
separation is preserved by safe (in the sense of Theorem
\ref{theorem-metalang-safety}) substitution of programs, assuming
disjointness of flips used in programs separated programs.
\begin{lemma}
  \label{lemma-scope}
  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and $\prog_1$
  covers $X$, then 
  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
  $\prog_1'$ such that $\vars(\prog_1') \cap \vars(\prog_2) = X$ and
  $\prog_1';\prog_2$ is safe.
\end{lemma}
\begin{proof}
  Suppose on the contrary that there exists $\prog_1'$ that is not
  separable from $\prog_2$. By results in
  \cite{barthe2019probabilistic}, this implies that there
  exists $\store^n$ and $\store^n$ for $0 \le n \le 1$ such that
  $\store^1_X = \store^{1'}_X$ where:
  $$
  \config{\store^0}{\prog_1;\prog_2} \redxs \config{\store^1}{\prog_2} \redxs
  \config{\store^2}{\varnothing}
  $$
  and:
  $$
  \config{\store^{0}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
  \config{\store^{2'}}{\varnothing}
  $$
  where $\store^2_{\vars(\prog_2)} \ne \store^{2'}_{\vars(\prog_2)}$, which is a
  contradiction.
%   This implies that there is some
%  $\store \in \mems(\vars(\prog_1';\prog_2))$ such that
%  $\store_{\vars{\prog_{1}'}} \in \runs{\prog_1'}$ and
%  \mems(\vars(\prog_1))$ with   
\end{proof}

We must also choose witnesses that allow us to generalize verification
of $\gNIMO$ preservation through gates. We need to consider separately
the definition of gates and their use. Our type discipline ensures that
gates can only be defined once, but it is also known that gate output
wire labels should only be used once \cite{nieminen2023breaking}, in which case we
say they are \emph{wired}. We return to this issue in more detail in
Section \ref{section-composition-copy}.
\begin{definition}
  We write $g \in \prog$ iff all views for $g$'s table are fully
  defined in $\prog$. We write $\wired(\prog)$ to denote the
  set of gates $g$ whose views occur on the rhs of an assignment
  in $\prog$.
  %$\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
  %and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
  %or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
  %and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
\end{definition}
Given the above definition, the needed result follows by Lemma \ref{lemma-scope} and definition
of $\gNIMO$:
\begin{lemma}
  \label{lemma-gnimo}
  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
  and
  $\gNIMO(\prog_1;\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ for
  $g \in \prog_2$,
  then $\gNIMO(\prog_1';\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
  for any $\prog_1'$ where $\prog_1';\prog_2$ is safe and
  $\vars(\prog_1') \cap \vars(\prog_2) = X$.
\end{lemma}

\subsubsection{Pre-~ and Post-Conditions}
\label{section-pre-post}
Now we can establish relevant pre-~and post-conditions for the
top-level YGC functions defined in Figure \ref{fig-ygc-gates}. We note
that in addition to the automated verification mechanisms described in
Section \ref{section-bruteforce}, type safety in $\metaprot$ plays an
important role, in part by ensuring that gates are uniquely defined
(as views).

Our strategy is to consider input encoding, internal gates, and
output decoding separately. To begin, we show that input encoding
establishes generates a program and wire values that initialize
the $\gNIMO$ invariant in the circuit. There is no precondition
witness needed for this result, which holds in any type-safe
program context since the variables used by $\ttt{encode}$
are all fresh. 
\begin{lemma}[Input Encoding]
  \label{lemma-encode}
  Given well-typed $\prog_1;E[\ttt{encode}(s_1,s_2)]$ where:
  $$
  \config{\prog_1}{E[\ttt{encode}(s_1,s_2)]} \redxs
  \config{\prog_1;\prog_2}{E[\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}]}
  $$
  Then:
  $$\gNIMO(\prog_1;\prog_2,v_1,s_1) \wedge \gNIMO(\prog_1;\prog_2,v_2,s_2)$$
\end{lemma}

\begin{proof}
  We have automatically verified
  $\gNIMO(\prog_2,v_1,\ttt{s1})$ and $\gNIMO(\prog_2,v_1,\ttt{s1})$
  where:
  $$
  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
  \config{\prog_2}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
  $$
  using the method described in Section \ref{section-bruteforce-tables}.
  Given arbitrary $\prog_1$ such that $\prog_1;\ttt{encode}(s_1,s_2)$
  is well-typed, hence $\prog_1;\prog_2$ is safe, 
  since $\ttt{encode}$ uses unique secrets and flips per gates and
  gates (views) are not redefined in safe programs,
  $(\iov(\prog_1) \cup \flips(\prog_1)) \cap (\iov(\prog_2) \cup
  \flips(\prog_2)) = \varnothing$. This implies 
  $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$, so the result follows
  by Lemma \ref{lemma-gnimo}.
\end{proof}

One nuance in our results for internal gates and for decoding is that the relevant
compositional units are the complete related subprotocols, including both
the garbler's construction of the gates and the evaluator's evaluation of
them. We therefore make the following definitions. However, we note that
our formal results are applicable to the YGC style where the garbler shares
the entire circuit prior to evaluation, rather than gate-by-gate during
evaluation. 
\begin{definition}
  \label{ygc-modules}
  We define a composable internal \ttt{and} gate sharing and evaluation function
  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
  $\ttt{decode}$, as follows. 
\begin{verbatimtab}
  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }

  decode(g, v) { garbledecode(g); evaldecode(v) }
\end{verbatimtab}
\end{definition}

Now we can demonstrate that our $\ttt{and}$ gate definition preserves
$\gNIMO$, and additionally that internale gates are probabilistically independent.
This latter fact is crucial to compositionality, and allows us for automatically
verify the circuit in isolation, with input encoding serving as an adequate input
witness for generalization. 
\begin{lemma}[Composition ($\eand$)]
  \label{lemma-compose}
Given well-typed $\prog_1;E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]$ where:
$$
\config{\prog_1}{E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
$$
If the following preconditions hold:
\begin{enumerate}
\item $(\{ g_1 \} \sqcup \{ g_2 \}) \cap \wired(\prog_1) = \varnothing$
\item $\gNIMO(\prog_1,v_1,g_1) \wedge \gNIMO(\prog_1,v_2,g_2)$ 
\end{enumerate}
then we have as postconditions:
\begin{enumerate}
  \item $\gNIMO(\prog_1;\prog_2,v,g)$
  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
\end{enumerate}
\end{lemma}
\begin{proof}
  We have automatically verified the postconditions (1) and (2) for the particular
  program $\prog_1;\prog_2$ resulting from execution of:
  $$e\quad \defeq\quad \ttt{let gs = encode(s1,s2) in andgg(0,s1,s2,gs.wv1,gs.wv2)}$$
  where:
  $$
  \config{\varnothing}{e} \redxs
  \config{\prog_1}{\ttt{andgg}(\ttt{0},\ttt{s1},\ttt{s2},v_1,v_2)} \redxs \config{\prog_1;\prog_2}{v}
  $$
  By Lemma \ref{lemma-encode} we have that
  $\gNIMO(\prog_1,v_1,\ttt{s1}) \wedge \gNIMO(\prog_1,v_2,\ttt{s2})$. 
  And since $\ttt{andgate}$ generates unique flips per gate, and well-typedness ensures
  uniqueness of gates, we can generalize the result to replace $\prog_1$ with arbitrary $\prog_1'$
  such that $\prog_1';\prog_2$ is safe,
  applying Lemma \ref{lemma-scope} to obtain postcondition (2) and Lemma \ref{lemma-gnimo}
  to obtain postcondition (1). 
\end{proof}
Finally, we can demonstrate that decoding correctly casts a garbled output into the
clear, and thus establishes $\NIMO$ for the garbled circuit that precedes it.
\begin{lemma}[Decoding]
  \label{lemma-decode}
Given well-typed $\prog_1;E[\ttt{decode}(g,v)]$ where
$$
\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
$$
If $\gNIMO(\prog_1,v,g)$, then $\NIMO(\prog_1;\prog_2;\eassign{\outv}{\be})$.
\end{lemma}
\begin{proof}
  We have automatically verified $
  \progd(\prog,\be)(\{ \itv \mapsto 1 \}) = 1$ where:
  $$\config{\varnothing}{\ttt{decode(g,\{ k = flip[2, gate:g.k]; p = flip[2,gate:g.p] \})}}
    \redxs \config{\prog}{\be}$$
    and also $
  \progd(\prog,\be)(\{ \itv \mapsto 0 \}) = 1$ where:
  $$\config{\varnothing}{\ttt{decode(g,\{ k = not flip[2, gate:g.k]; p = not flip[2,gate:g.p] \})}}
    \redxs \config{\prog}{\be}$$
  The result follows by definition of $\gNIMO$ and $\NIMO$.
\end{proof}

>>>>

\section{Component Certification and Composition}
\label{section-composition}

Using YGC as a case study, in this Section we will discuss
compositional properties that can be automatically verified on small
program components using techniques developed in Section
\ref{section-bruteforce}. These properties compose to yield $\NIMO$
and scale to arbitrary programs. In our metatheory we will express
these properties as pre-~and post-conditions on components, so our
general approach is in a proof assistant style.

@@

\subsection{Component Certification for $\NIMO$}
\label{section-composition-certification}

We can use the automated verification methods developed in Section
\ref{section-brute-force} to \emph{certify} YGC
components. Certification obtains properties that we can then use
higher-level reasoning-- inspired in part by techniques in
\cite{barthe2019probabilistic}-- to prove that circuits built with
certified components satisfy $\NIMO$ (Theorem \ref{theorem-ygc-NIMO})
and thus are passive secure.

Our method relies on a definition of compositional units comprising
complete related subprotocols-- i.e., both the garbler's
construction of the gates and the evaluator's evaluation of them. We
therefore make the following definitions. However, we note that our
formal results are applicable to the YGC style where the garbler
shares the entire circuit prior to evaluation.
\begin{definition}
  \label{ygc-modules}
  We define a composable internal \ttt{and} gate sharing and evaluation function
  called $\ttt{andgg}$, and an output decoder gate share and evaluation function called
  $\ttt{decode}$, as follows. 
\begin{verbatimtab}
  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }
  decode(g, v) { garbledecode(g); evaldecode(v) }
\end{verbatimtab}
\end{definition}

Our certification techniques are defined generally wrt implementations
of input encoding, decoding, and internal gates. In the following we
will refer to arbitrary decode, encode, and gate functions with the
restriction that any function in each of these categories has the same
valid type signature as $\ttt{decode}$, $\ttt{encode}$, and
$\ttt{andgg}$ respectively.

Certifications will demonstrate probabilistic independence of input
wire lables and garbled tables in a program's basic distribution, so
we need to consider dependencies of flips and gate views. Thus we make
the following definitions:
\begin{definition}
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$ as defined in Definition \ref{def-progd}. For any program $\prog =
  (\eassign{v_1}{\be_1};\ldots;\eassign{v_n}{\be_n})$, define
  $\vdefs(\prog) \defeq \{ v_1,\ldots,v_n \}$.
\end{definition}
We also need to formalize notions of probabilistic independence, aka
\emph{separation}, and correlation. We borrow the symbols $*$ and $\sim$
from \cite{barthe2019probabilistic} with the same denotations.
\begin{definition}
  We write $\vc{\pdf{X}}{x}{y}$ iff $\pdf{X}(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pdf{X}(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pdf{X}}{Y}{Z}$ iff for all
    $\store \in \mems(X \cup Y)$:
  $$(\margd{\pdf{X}}{Y \cup Z})(\store) =
  (\margd{\pdf{X}}{Y})(\store_Y) * (\margd{\pdf{X}}{Z})(\store_Z)$$ 
\end{definition}
It is also necessary to define notions of correlation with wire
labels, since wire values flow into and out of gates in either
positive of negative correlation with them.
\begin{definition}
  We write $\gc{\prog}{\{ \ttt{k = }\be_{k}; \ttt{p = }\be_{p} \}}{g}$ iff one of the
  following hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\vc{\progtt(\prog,\be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and} \quad
    \vc{\progtt(\prog,\be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
  \item $\vc{\progtt(\prog,\enot\ \be_k)}{\itv}{\ttt{flip[2,gate:}g\ttt{.k]}}\quad \text{and}\quad
    \vc{\progtt(\prog,\enot\ \be_p)}{\itv}{\ttt{flip[2,gate:}g\ttt{.p]}}$
  \end{enumerate}
\end{definition}
Thus, if a value is in correlation with a wire label, it is either
correlated with the wire label or with the inverse of both the key and
pointer bits. For subsequent discussion we make the following addition
to the codebase:
$$
\ttt{invert(\{ k = bk; p = bp \}) \{ \{ k = not bk; p = not bp \} \}}
$$
Gate certification establishes separation of input labels from the
gate table and output value, and correlation of the output with
the output wire label under any input wire value combination.
\begin{definition}[Gate Certification]
  \label{definition-gate-certification}
Let $\mathit{gate}$ be a gate function. For $1 \le i \le 4$  let:
  $$
  \config{\varnothing}{\ttt{v[0,out] := } e_i}\redxs\config{\prog_i}{\ttt{()}}
  $$
  where
  \begin{enumerate}[\hspace{5mm}$e_1 \defeq$]
    \item $\mathit{gate}\ttt{(c,a,b,owl(a),owl(b));}$
    \item $\mathit{gate}\ttt{(c,a,b,owl(a),invert(owl(b)));}$
    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),owl(b));}$
    \item $\mathit{gate}\ttt{(c,a,b,invert(owl(a)),invert(owl(b)))}$
  \end{enumerate}
  Define:
  $${\small X \defeq \ttt{\{ flip[2,gate:a.k], flip[2,gate:a.p], flip[2,gate:b.k], flip[2,gate:b.p] \}}}$$
  Then $\mathit{gate}$ is \emph{certified} iff for all $1 \le i \le 4$:
  $$
  \sep{\progtt(\prog_i)}{X}{\vdefs(\prog_i)} 
  \quad \text{and} \quad \gc{\progd(\prog_i)}{\vx{0}{out}}{\ttt{owl(c)}}
  $$
\end{definition}
Encode certification establishes correlation of the encoded input values with
the secret wire labels, and separation of client 2's secret from client 1's views. 
\begin{definition}[Encode Certification]
  Let $\mathit{encode}$ be an encoding function, and let:
  $$
  \config{\varnothing}{\ttt{encode(s1,s2)}} \redxs
  \config{\prog}{\ttt{\{wv1=}v_1\ttt{;wv2=}v_2\ttt{\}}}
  $$
  Then $\mathit{encode}$ is \emph{certified} iff each of the following conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\gc{\progd(\prog,v_1)}{\itv}{\ttt{owl(s1)}} \quad \text{and} \quad
    \gc{\progd(\prog,v_2)}{\itv}{\ttt{owl(s2)}}$
  \item $\sep{\progd(\prog,v_2)}{\{\ttt{s[2,s2]\}}}{(\vdefs(\prog) \cup \itv)}$
  \end{enumerate}
\end{definition}
Decode certification establishes separation of the input wire
label from the decoding tables under any input value condition. 
\begin{definition}[Decode Certification]
  Let $\mathit{decode}$ be a decoding function, and let:
  $$
  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{owl(c)})}
  \redxs \config{\prog_1}{\be_1}
  \quad \text{and} \quad
  \config{\varnothing}{\mathit{decode}(\ttt{c}, \ttt{invert(owl(c))})}
  \redxs \config{\prog_0}{\be_0}
  $$
  Then $\mathit{decode}$ is certified iff:
  $$
  \sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}
  $$
  %Then $\mathit{decode}$ is certified iff both of the following conditions hold:
  %\begin{enumerate}[\hspace{5mm}i.]
  %\item $\progd(\prog_1,\be_1)(\{ \itv \mapsto 1 \}) = 1$ and
  %  $\progd(\prog_0,\be_0)(\{ \itv \mapsto 0 \}) = 1$
  %\item $\sep{\progtt(\prog_i)}{\ttt{\{flip[2,gate:c.k],flip[2,gate:c.p]\}}}{\vdefs(\prog_i)}$
  %\end{enumerate}
\end{definition}
We have certified each of the YGC components detailed above. 
\begin{lemma}
  \label{lemma-certification}
  Each of $\ttt{decode}$, $\ttt{encode}$, and $\ttt{andgg}$ are certified.
\end{lemma}

\subsection{Correctness of Certified Components}
\label{section-composition-metatheory}

To demonstrate correctness of certification, we need to show that
isolated component certificates preserve relevant properties
when integrated into larger programs. This resonates with
\emph{framing} as developed in \cite{barthe2019probabilistic}.
In the following we will use $f$ to range over flips.
First, we show that if a flip $f$ is independent from the views
in a protocol (as is the case with input labels used to
encode a garbled table), then separation from views that
use $f$ (e.g., in an output label) is preserved. 
\begin{lemma}[Wire Framing]
  \label{lemma-wire-framing}
  If $\sep{\progtt(\prog)}{\{f\}}{\vdefs(\prog)}$ for flip $f$ and
  $\vars(\prog) \cap \vars(\prog') = \{ f \}$ then
  $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$.
\end{lemma}
Next, we show that output correlations (e.g., between output wire values and
labels) are preserved by substitution of correlated values for flips (e.g.,
between input wire values and labels).
\begin{lemma}[Substitution]
  \label{lemma-substitution}
  If $\vc{\progtt(\prog,\be)}{\itv}{f}$ and $\vc{\progtt(\prog',\be')}{\itv}{f'}$
  and $\sep{\progtt(\prog';\prog)}{\vdefs(\prog')}{\vdefs(\prog)}$
  then $\vc{\progtt(\prog';\prog,\be[\be'/f'])}{\itv}{f}$.
\end{lemma}
Our wire framing and substitution Lemmas then allow us to generalize
certification of gates and encoding to arbitrary circuit contexts,
assuming that outputs are not wired to multiple inputs \cite{tate2003garbled,nieminen2023breaking}.
\begin{lemma}[Gate Composition]
  \label{lemma-compose}
Given well-typed $\prog_1;E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]$ and certified $\mathit{gate}$ where:
$$
\config{\prog_1}{E[\mathit{gate}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
$$
If the following preconditions hold:
\begin{enumerate}
\item $\{ g_1 \} \cap \{ g_2 \} \cap \wired(\prog_1) = \varnothing$ and $g \not\in \prog_1$
\item $\gc{\progd(\prog_1)}{v_1}{\owl(g_1)}$ and $\gc{\progd(\prog_1)}{v_2}{\owl(g_2)}$
\end{enumerate}
then we have as postconditions:
\begin{enumerate}
  \item $\gc{\progd(\prog_1;\prog_2)}{v}{\owl(g_2)}$
  \item $\sep{\progd(\prog_1;\prog_2,v)}{\vdefs(\prog_1)}{\vdefs(\prog_2) \cup \{\itv\}}$
\end{enumerate}
\end{lemma}
\begin{proof}
  Given that $g_1$ and $g_2$ are distinct and not wired in $\prog_1$
  we are assured that $\ttt{owl}(g_1)$ and $\ttt{owl}(g_1)$ are in
  independent uniform distributions, and given that $g \not\in
  \prog_1$ we are assured that $\ttt{owl}(g)$ contains entirely fresh
  flips. Thus, precondition 2, Lemma \ref{lemma-wire-framing}, and
  Definition \ref{definition-gate-certification} imply postcondition
  2.  Furthermore, since gate certification guarantees that
  correlation with $\ttt{owl}(g)$ is guaranteed under all possible gate
  input wire values, precondition 2 and Lemma \ref{lemma-substitution}
  guarantee postcondition 2.
\end{proof}

\begin{lemma}[Decode Composition]
  \label{lemma-decode}
Given well-typed $\prog_1;E[\mathit{decode}(g,v)]$ and certified $\mathit{decode}$ where 
$$
\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
$$
If the following preconditions hold:
\begin{enumerate}
\item $g \not\in \cap \wired(\prog_1)$
\item $\gc{\progd(\prog_1)}{v}{\owl(g)}$
\end{enumerate}
then we have as postcondition:
\begin{enumerate}
  \item $\sep{\progd(\prog_1;\prog_2)}{\vdefs(\prog_1)}{\vdefs(\prog_2)}$
\end{enumerate}
\end{lemma}
Given the above, we can establish that any circuit built with certified
components is passive secure. 
\begin{theorem}
  \label{theorem-ygc-NIMO}
  If $e$ is a well-typed circuit definition using certified components
  where no gate output is wired more than once, and $\config{\varnothing}{e}
  \redxs \config{\prog}{\varnothing}$, then $\NIMO(\prog)$. 
\end{theorem}

\begin{proof}
  Let $\iov(\prog) = S \cup (V \cup \outv)$. Noting that any
  encoding necessarily uses entirely fresh variables by Definition and
  type safety, by applying Lemmas \ref{lemma-compose} and
  \ref{lemma-decode} and the frame rule of
  \cite{barthe2019probabilistic} we obtain:
  $$
  \sep{\progd(\prog)}{(S_{\{2\}}\cup\{\outv\})}{V}
  $$
  Noting that the garbler (client 2) has no views, this means that
  the garbler's secrets are independent of the evaluator's views. The
  result follows by Definition \ref{definition-NIMO}.
\end{proof}

### 2023-06-28 popl24/hyperprop.tex
checkpoint

\subsection{A Sound Hyperproperty for Passive Security}
\label{section-nimo}
The intuition that passive security means that the protocol releases
no more information than what is released by conditioning on the
public output and corrupt inputs suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories \emph{that can produce the same output} should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
  written $\NIMO(\prog)$, iff for all $H$ and $C$ with $|C|\le|H|$ and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  \begin{eqnarray*}
    & \store_1 =_C \store_2 \wedge
     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ \outv \}} =
     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ \outv \}}) \\
    & \implies \\
    & \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}
  \end{eqnarray*}
where $\iov(\prog) = (S,V)$.
\end{definition}

>>>>

\subsection{A Sound Hyperproperty for Passive Security}
\label{section-nimo}
The intuition that passive security means that the protocol releases
no more information than what is released by conditioning on the
public output and corrupt inputs suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories \emph{that can produce the same output} should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  \label{definition-NIMO}
  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
  written $\NIMO(\prog)$, iff for all $H$ and $C$ with $|C|\le|H|$ and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  \begin{eqnarray*}
    & \store_1 =_C \store_2 \wedge
     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ \outv \}} =
     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ \outv \}}) \\
    & \implies \\
    & \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}
  \end{eqnarray*}
where $\iov(\prog) = (S,V)$.
\end{definition}

### 2023-06-28 popl24/pmf.tex
checkpoint

In order to precisely characterize distributions of secrets and views
in programs, we define probability mass functions in a standard manner,
following the formulation in \ref{barthe2019probabilistic}. In this formulation, memories
denote conjunctions of variable assignments. Intuitively, given a
program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
\mems(\iov(\prog))$, when considering the execution of $\prog$
initialized with secrets $\store_S$, we are fundamentally concerned
with the joint probability of the assignments in $\store_V$
conditioned by the assignments $\store_S$. This is because
probabilistic dependence between honest secrets and corrupt views
\emph{is} information leakage.

>>>>

In order to precisely characterize distributions of secrets and views
in programs, we define probability mass functions in a standard manner,
following the formulation in \cite{barthe2019probabilistic}. In this formulation, memories
denote conjunctions of variable assignments. Intuitively, given a
program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
\mems(\iov(\prog))$, when considering the execution of $\prog$
initialized with secrets $\store_S$, we are fundamentally concerned
with the joint probability of the assignments in $\store_V$
conditioned by the assignments $\store_S$. This is because
probabilistic dependence between honest secrets and corrupt views
\emph{is} information leakage.
